Playing with PCAs

### 2024-03-15 3-reports/playing_with_pcas.Rmd
clusters

>>>>

## CLUSTERS

```{r}

library(cluster)
library(factoextra)

df = data_wide_PCA %>% filter(region == "CB") %>%   drop_na() %>% dplyr::select(where(is.numeric))
grp = data_wide_PCA %>% filter(region == "CB")%>% drop_na() %>% dplyr::select(!where(is.numeric))

fviz_nbclust(df, FUN = hcut, method = "wss")
fviz_nbclust(df, FUN = hcut, method = "silhouette")
# three clusters

d <- dist(df, method = "euclidean")
hc1 <- hclust(d, method = "complete" )
plot(hc1, cex = 0.6, hang = -1)
hc5 <- hclust(d, method = "ward.D2" )
sub_grp <- cutree(hc5, k = 3)
table(sub_grp)
plot(hc5, cex = 0.6)
rect.hclust(hc5, k = 3, border = 2:5)

df_cl <- mutate(df, cluster = sub_grp) %>% cbind(grp) %>% mutate(cluster = as.character(cluster))
count(df_cl,cluster)

```{r}
library(ggConvexHull)

pca_cluster = fit_pca_function(df_cl)

ggbiplot::ggbiplot(pca_cluster$pca_int, obs.scale = 1, var.scale = 1,
                   groups = as.character(pca_cluster$grp$cluster), 
                   ellipse = F, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_convexhull(aes(group = pca_cluster$grp$cluster,
                      fill = pca_cluster$grp$cluster), alpha = 0.2)+
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_cluster$grp$site,
                 color = groups))+ 
  # scale_color_manual(values = c("#16879C", "#BB281E" ))+
  #scale_shape_manual(values = c(21, 19))+
  #scale_shape_manual(values = c(21, 21, 19), name = "", guide = "none")+
  # xlim(-4,20)+
  # ylim(-8,8)+
  labs(fill="cluster",
       color = "cluster",
       shape = "region",
       title = "Overall PCA, both regions",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

```{r}
df_cl %>%
 # distinct(region, site, transect, cluster) %>% 
  ggplot(aes(y = cluster, x = site))+
  geom_point(aes(color = transect, shape = transect),
             size = 4, stroke = 1,
             position = position_jitterdodge(jitter.width = 0.0, jitter.height = 0.1, 
                                             dodge.width = 0.4))

devtools::install_github("liamgilbey/ggwaffle")
library(ggwaffle)

df_cl %>% 
  ggplot()+
  geom_waffle(aes(x = transect, y = site, fill = cluster))
